---
title: "3. Hypothesis Testing"
subtitle: "Adaptive Playmaking: Incorparating Prior Knowledge in Analyzing Football Passes Using Relational Event Modelling"
author: "Jonathan Koop"
date: "`r Sys.Date()`"
number-sections: true
format: html
---

This document provides the code for the Hypothesis Testing for the paper "Adaptive Playmaking: Incorparating Prior Knowledge in Analyzing Football Passes Using Relational Event Modelling". This is the first document in a series of documents that are used to analyze football passes using a relational event model. The documents are split into the following sections:

1.  Data Wrangling (`01_data_wrangling.qmd`)

2.  Sampling (`02_sampling.qmd`)

3.  **Hypothesis Testing (`03_hypothesis_testing.qmd`)**

This document relies on the functions provided in `functions/functions.Qmd` The sampled data is loaded from the `results/samples` folder. The results are saved in the `results` folder.

## Loading Relevant Packages

In order to run the code, the following packages need to be installed and loaded. The code below will install the packages if they are not already installed and load them.

```{r}
library(jsonlite)
library(dplyr)
library(purrr)
library(remstats)
library(remify)
library(MASS)
library(ggsoccer)
library(truncnorm)
library(extraDistr)
library(mvtnorm)
library(progress)
library(statmod)
library(abind)
library(ggplot2)
library(rvest)
library(stringr)
library(tidyr)
library(foreach)
library(doParallel)
library(xtable)
library(bain)
library(remstimate)

load("../../functions/functions.RData")
```

## Load Samples

```{r}
load("../../results/01_samples/results_gibbs_updated.RData")
```

# Hypothesis Testing Framework

## Using `bain`

```{r}
results_bain <- bain(
    x = results_gibbs_updated[[1]]$beta_mean,
    # mean of posterior
    "-b3<b2<b1<-b4;-b3<b1<b2<-b4;b2<-b3<b1<-b4;b1<-b3<b2<-b4",
    # hypotheses
    Sigma = results_gibbs_updated[[1]]$Sigma,
    # covariance matrix
    n = results_gibbs_updated[[1]]$M * results_gibbs_updated[[1]]$N
  )

results_bain
```

## Manual Computation

Due to the fact that no integration of a multivariate normal distribution with limits regarding the different variables is easily available, a sampling approach was taken. This is expected to be less accurate than the `bain` function as it is based on sampling from the prior and posterior distributions. Below, the results are computed.

```{r}
results_hypotheses <- bayesfactor(
  prior_mean = rep(0, 4),
  posterior_mean = results_gibbs_updated[[1]]$beta_mean,
  posterior_variance = results_gibbs_updated[[1]]$Sigma,
  n = results_gibbs_updated[[1]]$N,
  m = results_gibbs_updated[[1]]$M,
  n_samples = 10000000,
  method = "robust"
)

save(results_hypotheses, file = "../../results/02_hypotheses/results_hypotheses.RData")
```

And printed as a table, which is used in `report.qmd`.

```{r}
hypotheses_table <- print(xtable(results_hypotheses), include.rownames = FALSE)

# Save results as latex table
write(hypotheses_table, file = "../../output/02_tables/hypotheses_table.tex")
```

## Sensitivity Analysis

```{r}
results_hypotheses_sensitivity <- bayesfactor(
  prior_mean = rep(0, 4), # set prior mean to 0
  posterior_mean = results_gibbs_updated[[1]]$beta_mean, # set posterior mean
  posterior_variance = results_gibbs_updated[[1]]$Sigma, # set posterior variance
  b = results_bain$b, # set b to bain's b
  n_samples = 10000000, # set number of samples
  method = "bain" # set method to bain
)
```
